---
title: "ANOVA"
author: "Math 246"
date: "April 16, 2018"
output: 
  ioslides_presentation: 
    widescreen: yes
---

## ANOVA

Consider the facebook dataset. Does the avergae number of likes depend on what type of post you make?

Figure out the average number of likes for each of the types. What is the average number of likes for all posts? Distribution of the number of likes for each type? For all types?

## Example

```{r include=FALSE}
library(readr)
fb <- read_delim("../data/fb.csv", ";", escape_double = FALSE, trim_ws = TRUE)
library(tidyverse)
```

```{r}
fb %>% group_by(Type) %>% summarize(mean=mean(`Page total likes`))
```

## Thinking regression

Do a simple linear regression predicting likes from type of post. Get a summary and an anova table.

## Results {.smaller}

```{r}
modlinear <- fb %>% lm(`Page total likes`~Type, data=.)
summary(modlinear)
```
## ANOVA table

```{r}
anova(modlinear)
```


## Model

```{r}
mod1 <- fb %>% aov(`Page total likes`~Type, data=.)
summary(mod1)
```

## Meaning of ANOVA

```{r}
fb %>% ggplot(aes(x=Type, y = `Page total likes`))+geom_boxplot()
```

## Better yet

```{r message=FALSE, warning=FALSE}
fb %>% ggplot(aes(x=`Page total likes`))+geom_histogram()+facet_wrap(~Type, scales="free_y")
```


## Setup

Observation = Group mean + Error

Group mean = Grant mean + Group effect

Response = Grand mean + Group effect + Error term
$$Y = \mu + a_k + \epsilon$$

but

$$\mu_k = \mu + \alpha_k$$

## Plotted

```{r}
fb %>% ggplot(aes(x = Type, y = `Page total likes`))+geom_point()
```

## Back to the table: df

```{r}
mod1 <- fb %>% aov(`Page total likes`~Type, data=.)
summary(mod1)
```

## Hypothesis Test

$H_0: \alpha_1 = \alpha_2 = ... = \alpha_K = 0$

$H_A:$ at least one $\alpha_k \neq 0$

## What do these mean?

$\hat{\mu} = \bar{y}$

When is it true that $\hat{y} = \bar{y}$?

$\hat{\alpha}_k = \bar{y}_k - \hat{\mu} = \bar{y}_k - \bar{y}$

$\hat{\epsilon} = y - \hat{y} = y - \hat{\mu} - \hat{\alpha}_k = y - \bar{y}_k$

## So...

Estimated values:

$\hat{\mu} = \bar{y}$

$\hat{\alpha}_k = \bar{y}_k - \bar{y}$

$\hat{\epsilon} = y - \bar{y}_k$

## Assumptions

* Residuals have mean 0
* These normal populations have a common variance
  * Can use Levene’s test to check this
* Residuals are normally distributed (at each level)
  * (Kolmogorov–Smirnov test)
  * (Shapiro–Wilk test)
* Residuals are independent

## Resduals have mean 0

```{r}
mean(mod1$residuals)
```

```{r}
fb$residuals <- mod1$residuals
fb %>% group_by(Type) %>% summarise(mean=mean(residuals))
```

## Common variance

```{r}
fb %>% group_by(Type) %>% summarise(var = var(residuals))
```
```{r}
460521615/22378119
```
(2 would be better...)

## Normal residuals

```{r message=FALSE, warning=FALSE}
qplot(mod1$residuals)
```

## By type

```{r message=FALSE, warning=FALSE}
fb %>% ggplot(aes(x = residuals))+geom_histogram() + facet_wrap(~Type)
```

## Residuals vs fitted - same spread?

```{r}
qplot(mod1$fitted.values,mod1$residuals)
```

## How could we fix these broken assumptions?