---
title: "Data"
author: "Math 246"
date: "January 24, 2018"
output: ioslides_presentation
---

## NYT Article

## Where we're going
* Datacamp
* Manipulating data
* Graphing data
* Analysis, etc.

## Data Structure

* Open RStudio, we'll import the cdc dataset
* For each variable, what type is it?

## R Basics
* Importing data
* Assignment (<-)
* Picking variables ($)
* Functions (e.g. dim, as.factor)
* Packages
* factors
* data frames
* A note on `attach()` (why I don't think you should use it)

## Summaries and data wrangling

* There are many ways to do this (as there are for every task in R)
* We will use a package called `dplyr`
  * Part of the "tidyverse"
  * Your next homework will be to work through a unit on Datacamp
  
## dplyr
* tbls and tbl_df()
* pipe
* select
* mutate
* filter
* arrange
* summarize
* group_by  

## An example
`cdc %>% filter(gender=="f") %>% summarize(mean_height=mean(height))`

`cdc %>% filter(gender=="m") %>% summarize(mean_height=mean(height))`

can also do this other ways, but I think this is the most readable

`cdc %>% group_by(gender) %>% summarize(mean_height=mean(height))`

## An important feature of data sets

* *Generally* the order of columns and rows does not matter

## More "chapter 0"
* DATA = MODEL + ERROR
* Explanatory variables vs predictors
* Response variables
* Choose, fit, assess, use

## Notation
* $Y = \mu_i + \epsilon$
* $Y = \mu_1+\epsilon \sim N(\mu_1,\sigma_1)$
* Parameters are typically written with Greek characters
* Statistics are typically written with Roman/Latin characters, or with hats
* Residuals the difference between observed and predicted values

<!--

## A t-test with the cdc data set
* Reminder - what does a t-test do?

(We're not going to focus too much on this now, it's just an example)

```{r include=FALSE}
cdc <- read.csv("~/Dropbox/teaching/246page/data/cdc.csv")
require(dplyr)
```

> * A few ways to do this in order of how much I like them (also note - this matches the output style in the book)

## First

```{r}
t.test(height~gender, data=cdc)
```

## Second

```{r}
t.test(cdc$height~cdc$gender)
```

## Third

```{r}
t.test(cdc$height[cdc$gender=="m"],cdc$height[cdc$gender=="f"])
```

(this is actually useful if you have a factor variable with more than three levels)

## Putting together dplyr and t.test

```{r}
cdc %>% filter(age<70) %>% t.test(height~gender, data=.)
```

-->