---
title: "Logistic Regression"
author: "Math 246"
date: "March 22, 2018"
output: ioslides_presentation
---

## Why do we need this?

Load the hypothetical gre data from ucla dataset, and explore the data (summary).

(note: `xtabs` vs `count()`)

Say we want to predict whether someone is admitted to grad school... what could we do?

Q: What percent of the students were admitted to grad school?

## What kind of function (board work)

## Doing logistic regression with the grad school dataset

The `lm` command did linear regression, the `glm` command is the generalized linear model.

Format:

`model <- glm(y ~ x, data = mydata, family = "binomial")`

Why binomial?

## A Null model

Get a summary of a logistic regression predicting admit with no predictors (just use a 1 where )

## More models

Make a logistic model predicting admit from gre. Then make a second model predicting admit from gpa.

Write down both equations in both the logit form and the probability form. 

## The 50% mark

How can we figure out when we hit the 50% mark?

## New variable

Make a new column called `prestige` which is 1 if the school has rank 1 or 2, and 0 if it has rank 3 or 4.

Make a new logistic model predicting admit using `prestige`.

## Plotting {.smaller}

```{r include=FALSE}
require(readr)
require(tidyverse)
binary <- read_csv("https://stats.idre.ucla.edu/stat/data/binary.csv")
```

```{r}
binary %>% 
  ggplot(aes(x = gre, y = admit)) + 
  geom_point() +
  geom_smooth(method="glm", method.args = list(family = "binomial"), se=FALSE)
```

## Titanic data (titanic package)

```{r}
require(titanic)
d <- titanic_train
model1 <- d %>% glm(Survived~Age, data=.)

d$predicted <- predict(model1, d)
```

---

```{r message=FALSE, warning=FALSE}
d %>% 
  ggplot(aes(x=Age, y=Survived))+
  geom_point(color="red")+
  geom_smooth(method="glm", method.args = list(family = "binomial"), se=FALSE)+
  geom_point(color="green", aes(x=Age, y=predicted))
```

## What is the following going to do?

```{r, eval=FALSE}
d %>% 
  ggplot(aes(x=Age, y=Survived, colour=factor(Pclass)))+
  geom_point()+
  geom_smooth(method="glm", method.args = list(family = "binomial"), se=FALSE)+
  geom_point(aes(x=Age, y=predicted))
```

