---
title: "Multiple Regression"
author: "Math 246"
date: "February 21, 2018"
output: 
  ioslides_presentation: 
    widescreen: yes
---

## Back to the CDC dataset!! {.smaller}

```{r include=FALSE}
library(tidyverse)
library(readr)
cdc <- read_csv("../data/cdc.csv")
```

Predict height based on smoke100, how do we interpret the slope?

```{r}
cdc %>% lm(height~smoke100, data=.) %>% summary()
```

## Another model: predict height by exerany {.smaller}

```{r}
cdc %>% lm(height~exerany, data=.) %>% summary()
```

## Form of the model

Note: this is a model for the population (no hats!):

$$\mu_Y = \beta_0 + \beta_1 X_1 +\beta_2 X_2 + \ldots + \beta_k X_k$$

$$Y = \beta_0 + \beta_1 X_1 +\beta_2 X_2 + \ldots + \beta_k X_k + \epsilon$$

What's the difference?

## How to run it {.smaller}

What is the equation for the line of best fit?

```{r}
cdc %>% lm(height~exerany+smoke100, data=.) %>% summary()
```

## The new anova table

Check df - we now have 2 degrees taken up by the predictors, so we've got $n-2-1$ for the df of the residuals. (How many values are we predicting from our model?)

```{r}
cdc %>% lm(height~exerany+smoke100, data=.) %>% anova()
```

## The new standard error of the multiple regression model:

$$\hat{\sigma_\epsilon} = \sqrt{\frac{SSE}{df}} = \sqrt{\frac{SSE}{n-k-1}}$$

where we have $k$ predictors.

## More multivarite models  {.smaller}

How do we interpret these coefficients? What is the equation of the regression line?

```{r}
cdc %>% lm(height~age+exerany, data=.) %>% summary()
```

## One more model {.smaller}

```{r}
cdc %>% lm(height~age+weight, data=.) %>% summary()
```

## Plotting these things

## Assumptions

* Independence
* Constant variance for *any combination* of the predictors
