---
title: "Non-parametrics"
author: "Math 246"
date: "April 24, 2018"
output: 
  ioslides_presentation: 
    widescreen: yes
---

## Reminders

* What does a t-test do?
* What kinds of t-tests are there?
* What are the assumptions?

## T-test Assumptions

* Variable is continuous
* Random/Independent
* Data is normally distributed
* Sample size?
* Equal variance

## Relaxing the assumptions

* Continuous (not necessarily normal) distribution
  * Can actually use ordinal data
* Random/Independent

## The tests

| parametric                        | non-parametric                         |
|-----------------------------------|----------------------------------------|
| 1 sample t                        | sign test or Wilcoxon signed-rank test |
| 2 sample t (not paired)           | **Wilcoxon-Mann-Whitney**              |
| 2 sample paired t (matched pairs) | reduce to 1 sample                     |

## Wilcoxon-Mann-Whitney

* Toothpaste data set
* Why not use a t-test?

## Running the test by hand

$W_1 = n_1 n_2 + \dfrac{n_1(n_1+1)}{2}-R_1$

$W_2 = n_1 n_2 + \dfrac{n_2(n_2+1)}{2}-R_2$

## Toothpaste 1b

Try with the second dataset

## Toothpaste 2

## Making your own dataset

Create a small dataset which has a lot of overlap, compute $W_1$ and $W_2$.

Now do the same with a small dataset with no overlap.

What conclusions can you make about possible values of $W_1$ and $W_2$?

## More calculating

What is $W_1+W_2$ in general? Find a formula.

## Coding this

```{r include=FALSE}
require(ggplot2)
require(dplyr)
require(readr)
toothpaste1 <- read_csv("../data/toothpaste1.csv")
toothpaste1b <- read_csv("../data/toothpaste1b.csv")
toothpaste2 <- read_csv("../data/toothpaste2.csv")
toothpaste3 <- read_csv("../data/toothpaste3.csv")
```
```{r}
toothpaste2 %>% wilcox.test(score~brand, data=.)
```

Try 1b and 2

## A Question

Why did I say this would more more related to medians than means?

## Animal data

Download the animal data dataset. This includes information on animals and the length of time they're in a shelter before being adopted. What kinds of questions could we answer?

Consider some descriptive statistics / graphs.

## Animals

```{r include=FALSE}
AnimalData <- read_csv("../data/AnimalData.csv")
```

```{r}
AnimalData %>% ggplot(aes(x=Type, y=DaysInShelter, color=Type)) + 
  geom_boxplot() + geom_point()
```

## A significiant difference?

Is there a statistically significant difference in the length of time before adoption between cats and dogs?

## Cars

* The cars dataset has data cars from 1974.
* am is coded:
  * 0 = automatic
  * 1 = manual
* Is there a difference in mpg between automatic and manual transmission cars in 1974?

## Relationship to t-test

Consider the animal and cars analysis. Run each using a t-test. Would the t-test have been a reasonable alternative?